
<!-- Author: Jinming Cheng -->
<!-- Date: 17 July 2025 -->

<!-- Notes: Run knitr::knit_child() in a separate index Rmd file and set `sample_name` parameter in the environment -->

<!-- Usage: (R codes in index.Rmd):  -->
<!-- ``` -->
<!-- child_env <- new.env() -->
<!-- child_env$sample_name <- "Flex_Sample1" -->
<!-- res <- knitr::knit_child("01_03_Flex_Sample1_run_infercnv.Rmd", -->
<!--                          envir = child_env, quiet = TRUE) -->
<!-- ``` -->

## `r sample_name`

```{r,include=FALSE}
eval_infercnv = TRUE
eval_heatmap4 = TRUE
eval_save_rdata = TRUE
```

```{r, include=FALSE}
load("results/00_prepare.rdata")
```

Parameters
```{r}
data_for_infercnv = paste0(rds_res_dir,"/","dge_for_infercnv_",sample_name,".rds")
out_dir  = paste0(result_dir,"/","infercnv_res","/",sample_name)
out_rdata_file =  paste0(out_dir,"/","infercnv_out.rdata") 
```

```{r}
if (!dir.exists(out_dir)) {
  dir.create(out_dir, recursive = TRUE)
}
```


Read in data

```{r}
library(infercnv)
```

```{r}
dge_for_infercnv = readRDS(file = data_for_infercnv)

# remove the first column in gene table if it is gene symbol, 
# because genes are already in the rownames, and infercnv use 'chr' as the first column
# dge_for_infercnv$genes = dge_for_infercnv$genes[,c("chr","start","end")] 
```


sample information for infercnv (only the first column will be used)
```{r}
head(dge_for_infercnv$samples)
```


gene information for infercnv
```{r}
head(dge_for_infercnv$genes)
```


reference group names
```{r}
dge_for_infercnv$ref_groups
```


Create infercnv object

<!-- (only the first column in dge_for_infercnv$samples will be used by infercnv) -->
```{r, results='hide', eval=eval_infercnv}
infercnv_obj <- CreateInfercnvObject(
  raw_counts_matrix = dge_for_infercnv$counts,    # A count matrix
  annotations_file  = dge_for_infercnv$samples,   # A data frame with sample annotation 
  gene_order_file   = dge_for_infercnv$genes ,    # A data frame with gene annotation with chr, coords
  ref_group_names   = dge_for_infercnv$ref_groups # A vector with reference group names
)
```


Run infercnv
```{r, results='hide', eval=eval_infercnv}
infercnv_output <-
  infercnv::run(infercnv_obj,
                # without this flag, looks for pre-computed data in output directory.
                resume_mode = FALSE,
                cutoff = 0.1,
                # cutoff=1 works well for Smart-seq2, and cutoff=0.1 works well for 10x Genomics
                out_dir = out_dir,
                cluster_by_groups = TRUE,
                denoise = TRUE,
                # Ignore HMM-based estimates of CNVs for now.
                # This takes quite a lot of memory and is probably a Milton job.
                HMM = FALSE
  )
```


Draw heatmaps using ComplexHeatmap
```{r}
filename_infercnv_obj = paste0(out_dir,"/","run.final.infercnv_obj")
data_for_plot_cnv = readRDS(filename_infercnv_obj)
```

```{r}
library(ComplexHeatmap)
```

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

### Figure 2B

Heatmap with chromosome annotation on the left and legend on the right (use list)

<!-- notes: for each sub heatmap -->
<!-- column_split = list_ordered_groups[[i]]$sample -->

```{r,results='hide',eval=eval_heatmap4,include=FALSE}

ordered_groups = do.call(c, list(data_for_plot_cnv@observation_grouped_cell_indices,
                                 data_for_plot_cnv@reference_grouped_cell_indices) )

ordered_groups = cbind.data.frame( group = rep(names(ordered_groups),
                                    c(lapply(ordered_groups, length))),
                        position = do.call(rbind,
                                           lapply(ordered_groups, data.frame) )[,1] )

# use ordered levels in the raw dge data
ordered_groups$group = factor(ordered_groups$group,
                              levels = levels(dge_for_infercnv$samples$new_group) )


data_for_heatmap = data_for_plot_cnv@expr.data[, ordered_groups$position]

# add sample information from dge object # ** dge_for_infercnv ** #
ordered_groups$sample = dge_for_infercnv$samples[colnames(data_for_heatmap),"group"]


# colors for group used in infercnv
# my_colors_col = my_colors_20
my_colors_col =  c(my_colors_20, rev(viridis::plasma(10)) )  # add more colors on 06 Feb 2024
color_groups_col = my_colors_col[1:nlevels(ordered_groups$group)]
names(color_groups_col) = levels(ordered_groups$group)

# colors for subsetting data
ordered_groups$group = droplevels(ordered_groups$group)
color_groups_col = color_groups_col[levels(ordered_groups$group)]

# colors for sample information from dge data
# my_colors_col = RColorBrewer::brewer.pal(8,name = "Accent")
my_colors_col = col.p
color_samples_col = my_colors_col[1:nlevels(ordered_groups$sample)]
names(color_samples_col) = levels(ordered_groups$sample)


# column_ha = HeatmapAnnotation( group  = ordered_groups$group,
#                                sample = ordered_groups$sample,
#                                col = list(group  = color_groups_col,
#                                           sample = color_samples_col),
#                                show_legend = TRUE, # show legend
#                                show_annotation_name = FALSE )

list_data_for_heatmap = list()
list_ordered_groups = list()
list_column_ha = list()
list_ht_raw = list()

for (i in levels(ordered_groups$group) ) {
  keep_columns <- (ordered_groups$group==i)
  list_data_for_heatmap[[i]] = data_for_heatmap[,keep_columns]
  list_ordered_groups[[i]] = ordered_groups[ordered_groups$group==i,]

  list_column_ha[[i]] = HeatmapAnnotation( group  = list_ordered_groups[[i]]$group,
                                 sample = list_ordered_groups[[i]]$sample,
                               col = list(group  = color_groups_col,
                                          sample = color_samples_col),
                               show_legend = TRUE, # show legend
                               show_annotation_name = FALSE )

ht =  Heatmap(matrix = list_data_for_heatmap[[i]],

        width = 7,

        # column anno #
        top_annotation =  list_column_ha[[i]],
        cluster_columns = TRUE,
        show_column_dend = FALSE, # show denrogram or not
        show_parent_dend_line = FALSE, # remove dashed line of parent dendrogram
        show_heatmap_legend  = FALSE,
        column_split = list_ordered_groups[[i]]$sample,
        column_title = NULL,    # show column title or not (level names in `column_split`)
        column_gap = unit(0, "mm"), # column split gap
        # column_title_rot = 0,

        # row anno #
        cluster_rows = FALSE,
        row_split = data_for_plot_cnv@gene_order$chr,
        # row_title = NULL,      # show row title or not (level names in `row_split`)
        row_title_rot = 0,

        use_raster = TRUE,

        show_column_names = FALSE,
        show_row_names = FALSE)


 list_ht_raw[[i]] = ht

 if(i==levels(ordered_groups$group)[1]){
   list_ht = ht
 }else{
   list_ht = list_ht + ht
 }

}

output_filename = paste0(out_dir,"/","ComplexHeatmap_infercnv_with_sampleInfo_equal_width.png")

date()

png(filename = output_filename,
    width = 8,height = 8,units = "in",res = 300)


ComplexHeatmap::draw(list_ht,ht_gap = unit(1, "mm"))

dev.off()

date()
```


```{r,include=FALSE,eval=eval_heatmap4,include=FALSE}
filename_rds = paste0(out_dir,"/","heatmap_list_combined.rds")
saveRDS(list_ht,file = filename_rds)

filename_rds = paste0(out_dir,"/","heatmap_list_raw.rds")
saveRDS(list_ht_raw,file = filename_rds)
```


```{r,include=FALSE}
output_filename = paste0(out_dir,"/","ComplexHeatmap_infercnv_with_sampleInfo_equal_width.png")
```

inferCNV heatmap using ComplexHeatmap (chr on the left, legend on the right with sample information)
```{r, echo = FALSE, out.width='100%', fig.align='center'}
my_output_filename = output_filename
knitr::include_graphics(my_output_filename)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

<!-- Save rdata -->
```{r,eval=eval_save_rdata,include=FALSE}
save.image(file = out_rdata_file)
```



calculate instability score
```{r}
instability_score = apply((data_for_heatmap-1)**2, 2, mean)
range(instability_score)
```


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

### Figure 2C

Boxplot of instability score
```{r,include=FALSE}
data_for_plot = cbind.data.frame(instability_score, ordered_groups)

my_colors =  c(my_colors_20, 
               rev(viridis::plasma(10)) ) 
p1 <- ggplot(data = data_for_plot,
             aes(x=group, 
                 y=instability_score, color=group ) ) + 
  geom_boxplot(show.legend=FALSE) + 
  scale_color_manual(values=my_colors ) + 
  labs(x="", y="inferCNV instability score",title = sample_name) +
  # coord_cartesian(ylim=c(0,0.02) ) +
  my_theme_ggplot + my_theme_facet + Seurat::RotatedAxis()


n_scale = nlevels(ordered_groups$group)/10

my_output_filename = paste0(out_dir,"/",
                            "Boxplot_infercnv_instability_score",".png")

png(my_output_filename,
    height = 4*1,
    width = 4*1*n_scale,
    units = "in",res = 300)
p1
dev.off()
```

```{r, echo = FALSE, out.width='60%', fig.align='center'}
knitr::include_graphics(my_output_filename)
```


<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

\pagebreak
