
<!-- Author: Jinming Cheng -->
<!-- Date: 17 July 2025 -->

# Flex single cell data

## Sample1

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

```{r,include=FALSE}
eval_Seurat = FALSE
# eval_fina_all_markers = FALSE
eval_save_data_seurat_cell_type = TRUE
# eval_save_rdata = TRUE
```


```{r, include=FALSE}
load("results/00_prepare.rdata")

library(Seurat)
library(ggplot2)
```

### Read in data

Read in raw count matrix
```{r}
filename_h5 = paste0(data_dir,"/",
                     "Flex/GSM7782698_count_raw_feature_bc_matrix.h5")

h5_data = Read10X_h5(filename_h5)
```


Number of genes and cells (not filtered)
```{r}
dim(h5_data)
```


Create seurat object
```{r}
data_seurat_raw <- CreateSeuratObject(counts = h5_data, 
                                      project = "Sample1", 
                                      min.cells = 3, 
                                      min.features = 0)
```

```{r}
dim(data_seurat_raw)
```


Filter out gene names starting with "DEPRECATED-"
```{r}
n_rows_to_filter = grep("DEPRECATED-",rownames(data_seurat_raw))
```

Remove deprecated genes
```{r}
data_seurat_raw = data_seurat_raw[-n_rows_to_filter,]
```


```{r}
dim(data_seurat_raw)
```


Mitochondria read percentage
```{r}
data_seurat_raw[["percent.mt"]] <- PercentageFeatureSet(data_seurat_raw, 
                                                        pattern = "^MT-")
```


Prefix of output file names
```{r}
sample_name = c("Flex_Sample1")
one_group <- sample_name
```

Prefix of figures
```{r outputPrefix}
my_output_prefix_figure = paste0(my_output_prefix,"figures/",sample_name,"/")
if(!dir.exists(my_output_prefix_figure)){dir.create(my_output_prefix_figure,recursive = TRUE)}
```

### Quality control

read in cell type annotation from 10x
```{r}
filename_csv = paste0(data_dir,"/","Cell_type_annotation/",
                      "Cell_type_annotation_scFFPE_seq.csv")
cell_type_anno_10x = read.csv(filename_csv)
rownames(cell_type_anno_10x) = cell_type_anno_10x$Barcode
dim(cell_type_anno_10x)
```

check cell existence (should be true)
```{r}
all(rownames(cell_type_anno_10x) %in% colnames(data_seurat_raw) )
```


Use annotated cells from 10x 
```{r}
data_seurat = data_seurat_raw[,rownames(cell_type_anno_10x)]
```


N cells and genes after QC
```{r}
dim(data_seurat)
```

save filtered data as dge object for inferCNV analysis 
```{r}
df_genes = as.data.frame(rownames(data_seurat))
colnames(df_genes) = "Symbol"
dge_data = edgeR::DGEList(counts = as.matrix( GetAssay(data_seurat,assay = "RNA")@counts ),
                          samples = data_seurat@meta.data,
                          genes = df_genes)
dge_data$samples$group = sample_name
```


<!-- save dge data into rds for inferCNV analysis -->
```{r,include=FALSE}
dge_data_dir=paste0(data_dir,"/","dge_data")
filename_rds= paste0(dge_data_dir,"/","dge_",sample_name,".rds")
saveRDS(dge_data, file =filename_rds)
filename_rds = "temp"
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

### Seurat analysis

set parameters 
```{r}
n_var_genes <- 1000
n_dims      <- 30
resolution  <- 0.2
```

Run Seurat pipeline and predict doublets
```{r, eval=eval_Seurat}
data_seurat = SeuratPipelineAfterQC(data_seurat = data_seurat,
                                    n_var_genes = n_var_genes,
                                    n_dims = n_dims,
                                    resolution = resolution,
                                    double_rate = 0.1 )

```


```{r}
dim(data_seurat)
```


Backup raw clusters
```{r}
data_seurat$seurat_clusters.raw <- data_seurat$seurat_clusters
```


Find sub clusters for cluster 7 (epithelial). Decide to run FindSubCluster after looking the downstream feature plots of marker genes.
```{r}
data_seurat =  FindSubCluster(data_seurat,
                              graph.name = "RNA_snn",
                              cluster = "7",
                              resolution = 0.1)
```

Assign sub clusters to seurat clusters and active.ident
```{r}
ordered_numer = sort(table(data_seurat$sub.cluster),decreasing = TRUE)
data_seurat$sub.cluster = factor(data_seurat$sub.cluster,
                                 levels = names(ordered_numer) )

data_seurat$seurat_clusters = factor(data_seurat$sub.cluster,
                                     levels = names(ordered_numer),
                                     labels = 0:(length(ordered_numer)-1) )
data_seurat@active.ident <- data_seurat$seurat_clusters
```


Find sub clusters for new cluster 7 (B and Plasma). Decide to run FindSubCluster after looking the downstream feature plots of marker genes.
```{r}
data_seurat =  FindSubCluster(data_seurat,
                              graph.name = "RNA_snn",
                              cluster = "7",
                              resolution = 0.1)
```

Assign sub clusters to seurat clusters and active.ident
```{r}
ordered_numer = sort(table(data_seurat$sub.cluster),decreasing = TRUE)
data_seurat$sub.cluster = factor(data_seurat$sub.cluster,
                                 levels = names(ordered_numer) )

data_seurat$seurat_clusters = factor(data_seurat$sub.cluster,
                                     levels = names(ordered_numer),
                                     labels = 0:(length(ordered_numer)-1) )
data_seurat@active.ident <- data_seurat$seurat_clusters
```

Find sub clusters for new cluster 7 (pericyte and endothelial). Decide to run FindSubCluster after looking the downstream feature plots of marker genes.
```{r}
data_seurat =  FindSubCluster(data_seurat,
                              graph.name = "RNA_snn",
                              cluster = "7",
                              resolution = 0.05)
```

Assign sub clusters to seurat clusters and active.ident
```{r}
ordered_numer = sort(table(data_seurat$sub.cluster),decreasing = TRUE)
data_seurat$sub.cluster = factor(data_seurat$sub.cluster,
                                 levels = names(ordered_numer) )

data_seurat$seurat_clusters = factor(data_seurat$sub.cluster,
                                     levels = names(ordered_numer),
                                     labels = 0:(length(ordered_numer)-1) )
data_seurat@active.ident <- data_seurat$seurat_clusters
```


Predict cell cycle phase
```{r}
s.genes <- subset(human_cc_genes,Phase=="S")$Symbol
g2m.genes <- subset(human_cc_genes,Phase=="G2M")$Symbol
data_seurat <- CellCycleScoring(data_seurat,  
                                s.features = s.genes,g2m.features = g2m.genes)
```


<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

<!-- figures for the 2025. BMC. paper -->

#### Figure 2A

DimPlot by clusters
```{r, fig.width=4*2*1, fig.height = 3*1*1, results='hide',include=FALSE}
p <- list()
my_group_by <- c("seurat_clusters")
my_colors  <- my_colors_20

for (i in 1:length(my_group_by)) {
  p[[i]] <- DimPlot(data_seurat,
                    pt.size=0.1,
                    reduction="umap",
                    group.by=my_group_by[i],
                    shuffle=TRUE,
                    cols=my_colors) +
    my_theme_ggplot +
    ggplot2::labs(title="")
}

my_output_filename = paste0(my_output_prefix_figure,
                            one_group,"_",
                            "UMAP_plot_clusters",
                            ".png")

png(my_output_filename,
    width = 4*1*1.8, height = 3*1*1.8, units = "in",res = 300)
patchwork::wrap_plots(p, nrow=1, byrow=FALSE)
dev.off()
```

```{r, echo = FALSE, out.width='60%', fig.align='center'}
knitr::include_graphics(my_output_filename)
```


#### Figure 2D

Dimplot by cell types
```{r, results='hide',include=FALSE}
data_seurat$cell_type_label = data_seurat$seurat_clusters
levels(data_seurat$cell_type_label) = c("Macrophage",
                                        "T",
                                        "Fibroblast",
                                        "Tumor",  # "Tumor_ML"
                                        "Tumor",  
                                        "DCIS",   # "DCIS_ML"
                                        "DCIS",
                                        "Basal",
                                        "Endothelial",
                                        "B",
                                        "LP",
                                        "Plasma",
                                        "Pericyte",
                                        "Lymph.vessel",
                                        "pDC",
                                        "Mast",
                                        "ML",
                                        "Adipocyte",
                                        "Tumor"
                                        )

nlevels(data_seurat$cell_type_label)

ordered_numer = sort(table(data_seurat$cell_type_label),decreasing = TRUE)
data_seurat$cell_type_label = factor(data_seurat$cell_type_label,
                                     levels = names(ordered_numer) )


xlim_range = range(data_seurat@reductions$umap@cell.embeddings[,1])
my_xlim = c( floor(xlim_range[1]/5)*5,
            ceiling(xlim_range[2]/5)*5 )


my_plot = Seurat::DimPlot(data_seurat,
                pt.size = 0.1,
                reduction = "umap",
                label = FALSE,
                group.by = "cell_type_label",
                shuffle = TRUE,
                cols =  my_colors_20)  +
  my_theme_ggplot +
  ggplot2::labs(title = one_group) +
  Seurat::NoLegend() +
  ggplot2::scale_x_continuous(limits = my_xlim )


my_plot = Seurat::LabelClusters(my_plot,
                                id = 'cell_type_label',
                                size = 4,
                                repel = TRUE,
                                nudge_x = c( 4,-3,-3, 0,-4, 0,
                                             2, 0,-3, 0, 3, 3,
                                             3, 2, 4, 5 ),
                                nudge_y = c( 2,-4, 0, 4,-2, 2,
                                             3,-2, 0, 2, 0,-1,
                                             0,-2, 0,-1 )  )

# nudge_x and nudge_y orders
# 1. Endothelial, 2. DCIS, 3. T, 4. Macrophage, 5. Tumor, 6. Basal,
# 7. Fibroblast, 8. LP, 9. B, 10. Pericyte, 11. Plasma, 12. Mast,
# 13. pDC, 14. ML, 15. Adipocyte, 16. Lymph.vessel,


my_output_filename = paste0(my_output_prefix_figure,one_group,"_",
                            "DimPlot_cell_type_annotation",".png")
png(my_output_filename,width = 8*0.7,height = 6*0.7,units = "in",res = 300)

my_plot

dev.off()
```

```{r, echo = FALSE, out.width='60%', fig.align='center'}
knitr::include_graphics(my_output_filename)
```


#### Figure 2E

```{r}
my_genes_list = list("Epithelial" = c("EPCAM"), #  "ITGA6"
                     "Stromal" = c("IGFBP7"),
                     "Cycling" = c("MKI67","HMGB2"),
                     "Basal" = c("KRT14", "ACTA2"),
                     "LP" = c("ELF5", "MMP7"),
                     "ML" = c("PRLR", "AREG"),
                     "Tumor" = c("PTGDS","COL6A2"),  # MpBC markers for one subtype
                     "TumorER" = c("SCGB2A1"),      # tumor ML (ER)
                     "ML2" = c("ABCC11"),      # ABCC11 is expressed in both tumor ML (HER2 tumor) and normal ML
                     "DCIS" = c("CEACAM6"),      # tumor ML (HER2 DCIS)
                     "Fibroblast" = c("DCN","LUM", "PDGFRA", "COL1A2"), #PI16
                     "Endothelial" = c("PECAM1", "VWF","CD36", "CD34"),
                     "Pericyte" = c("RGS5", "MCAM"),
                     "Adipocyte" = c("ADIPOQ","FABP4"),
                     # "Schwann" = c("MPZ", "MBP"),
                     "T" = c("CD3G", "CD3D","CD4","CD8A"),
                     "NK" = c("NKG7", "NCR1"),
                     "B" = c("MS4A1", "CD79A"),
                     "Plasma" = c("JCHAIN", "IGKC"),
                     "Macrophage" = c("C1QA","CD68"),
                     "Dendritic"  = c("CD1C","FCER1A","IDO1","CST3","ITGAX","IRF8"),
                     "pDC" = c("GZMB", "IRF7"),
                     # FDCSP gene is a marker for follicular dendritic cells (FDCs)
                     # "Mast" = c("TPSB2", "TPSAB1"),
                     "Mast" = c("CTSG","MS4A2"),
                     "Lymph.vessel" = c("LYVE1", "PROX1"),
                     # "Muscle" = c("DES", "MYOG"),
                     "Inflam.Macro" = c("S100A8","S100A9") # could also be dendritic, 2022.11.01

)

my_genes = do.call(c, my_genes_list)
```


DotPlot of my genes by cell types
```{r, results='hide',include=FALSE}
my_genes_list_df = cbind.data.frame(Symbol = unlist(my_genes_list),
                                   cell_type = rep(names(my_genes_list),
                                                lapply(my_genes_list, length)))

anno_cell_type = strsplit(levels(data_seurat$cell_type_label),split = "[,_] *")
names(anno_cell_type) = levels(data_seurat$cell_type_label)
anno_cell_type = data.frame("cell_type" = unlist(anno_cell_type),
                            "cell_anno" = rep(names(anno_cell_type),
                                              lapply(anno_cell_type,length)) )

my_genes_ordered = subset(my_genes_list_df,
                          cell_type %in% anno_cell_type$cell_type )

diff_cell_types = setdiff(levels(data_seurat$cell_type_label),
                          unique(my_genes_ordered$cell_type))


data_seurat$cell_type_ordered = data_seurat$cell_type_label
data_seurat$cell_type_ordered =
  factor(data_seurat$cell_type_ordered,
         levels = c(unique(my_genes_ordered$cell_type), diff_cell_types) )

# nlevel should be the same
nlevels(data_seurat$cell_type_ordered) == nlevels(data_seurat$cell_type_label)
my_genes_ordered = my_genes_ordered$Symbol

my_genes_ordered = unique(c("EPCAM",
                            "ESR1","PGR","ERBB2",
                            my_genes_ordered))

my_output_filename = paste0(my_output_prefix_figure,
                            one_group,"_",
                            "DotPlot_marker_genes_cell_type",".png")
png(my_output_filename,
    height = ceiling(2*length(my_genes_ordered)/7)+2,
    width = ceiling(2*nlevels(data_seurat$cell_type_label)/4)+1,
    units = "in",res = 300)

Seurat::DotPlot(object = data_seurat,
                group.by = "cell_type_ordered",
                features = my_genes_ordered ) +
  my_theme_ggplot + Seurat::RotatedAxis() +
  ggplot2::labs(y = "Cell type", x = "Genes")  +
  ggplot2::coord_flip()

dev.off()
```

```{r, echo = FALSE, out.width='60%', fig.align='center'}
knitr::include_graphics(my_output_filename)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- Save rds of one sample -->
```{r, include=FALSE,eval=eval_save_data_seurat_cell_type}
filename_rds = paste0(rds_res_dir,"/", "data_seurat_cell_type","_", one_group,".rds")
saveRDS(data_seurat, file = filename_rds)
```


\pagebreak
