
<!-- Author: Jinming Cheng -->
<!-- Date: 17 July 2025 -->

<!-- Notes: The "Rep1_Xenium_v2_list_selected_cells_from_sampling.rds" file  -->
<!-- in this Rmd file is generated by the sampling process in -->
<!-- 02_04_Xenium_Sample1_multiple_runs_for_runtime_comparing.Rmd -->

## Reference for multiple runs

To investigate relationship of the running time of the reference-based methods with the number of cells of the reference data, we split the reference single-cell datasets into smaller subsets and created smaller reference datasets to run the pipeline multiple times.

```{r,include=FALSE}
eval_prepare_azimuth_ref = FALSE
eval_prepare_RCTD_reference = FALSE
```


### Azimuth reference (Flex_Sample1)

The reference is prepared for multiple runs of Rep1 Xenium v2 to compare running time using different number of cells in the reference.

```{r,include=FALSE}
load("results/00_prepare.rdata")
```


Load packages
```{r loadPackage}
library(Seurat)
library(ggplot2)
library(edgeR)
library(SingleR)
library(Azimuth)
library(spacexr)
```


read in scRNASeq reference data
```{r}
filename_rds = paste0(rds_res_dir,"/", "data_seurat_cell_type","_", "Flex_Sample1",".rds")
data_seurat = readRDS(file = filename_rds)
```

read in list_selected_cells 
```{r,include=FALSE}
filename_rds = paste0(rds_res_dir,"/",  "Rep1_Xenium_v2","_",
                      "list_selected_cells_from_sampling",".rds")
list_selected_cells <- readRDS(file = filename_rds)
```

```{r}
names(list_selected_cells)
```


keep singlets as reference
```{r}
data_seurat_singlet = data_seurat[,data_seurat$db_type=="singlet"]
```


prepare azimuth reference for each reference subset
```{r,results='hide',eval=eval_prepare_azimuth_ref}
for (i in names(list_selected_cells) ) {
  
  # create azimuth reference dir
  ref_azimuth_dir = paste0(my_output_prefix,"ref_azimuth/Flex_Sample1","_",i)
  if(!dir.exists(ref_azimuth_dir)){dir.create(ref_azimuth_dir,recursive = TRUE)}
  
  # keep a subset of cells
  cells_kept = list_selected_cells[[i]]$cell_id
  data_seurat_sub = data_seurat_singlet[,cells_kept]
  
  # Run umap with reurn.model and normalize data using SCT method for azimuth
  data_seurat_sub = RunUMAP(data_seurat_sub, dims=1:30,return.model = TRUE)
  data_seurat_sub = SCTransform(data_seurat_sub)
  
  # Prepare azimuth reference data
  ref_azimuth = AzimuthReference(data_seurat_sub,
                                 refUMAP = "umap",
                                 refDR = "pca",
                                 metadata = c("cell_type_label") )
  saveRDS(ref_azimuth,file = paste0(ref_azimuth_dir,"/","ref.Rds"))
  SaveAnnoyIndex(ref_azimuth[["refdr.annoy.neighbors"]], 
                 file = paste0(ref_azimuth_dir,"/","idx.annoy"))
}
```


### RCTD reference (Flex_Sample1)

```{r,include=FALSE}
library(spacexr)
```


prepare RCTD reference for each reference subset
```{r,results='hide',eval=eval_prepare_RCTD_reference}
for (i in names(list_selected_cells) ) {
    
  # create RCTD reference dir
  ref_RCTD_dir = paste0(my_output_prefix,"ref_RCTD/Flex_Sample1","_",i)
  if(!dir.exists(ref_RCTD_dir)){dir.create(ref_RCTD_dir,recursive = TRUE)}
  
  # keep a subset of cells
  cells_kept = list_selected_cells[[i]]$cell_id
  data_seurat_sub = data_seurat_singlet[,cells_kept]
  
  # Prepare reference
  counts <- GetAssayData(data_seurat_sub, assay = "RNA", slot = "counts")
  cell_type <- as.factor(data_seurat_sub$cell_type_label)
  names(cell_type) <- colnames(data_seurat_sub)
  nUMI <- data_seurat_sub$nCount_RNA
  ref_RCTD <- Reference(counts, cell_type, nUMI)
  
  # save ref
  saveRDS(ref_RCTD,file = paste0(ref_RCTD_dir,"/","reference.rds"))

}
```

\pagebreak
