
<!-- Author: Jinming Cheng -->
<!-- Date: 17 July 2025 -->

# Sample2 (v2)

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

```{r,include=FALSE}
eval_read_in_raw_data = FALSE
eval_Seurat = FALSE
eval_fina_all_markers = FALSE
eval_predict_cell_type_SingleR = TRUE 
eval_predict_cell_type_azimuth = TRUE
eval_predict_cell_type_scmapCell = TRUE
eval_predict_cell_type_RCTD = TRUE
eval_predict_cell_type_scPred = TRUE
eval_save_xenium_obj = TRUE
eval_save_rdata = TRUE
```


## 10x Xenium

### Read in data

```{r,include=FALSE}
load("results/00_prepare.rdata")
```

Load packages
```{r loadPackage}
library(Seurat)
library(ggplot2)
library(edgeR)
library(SingleR)
library(Azimuth)
library(scmap)
library(spacexr)
library(scPred)
```

Prefix of output file names
```{r sampleName}
sample_name = "Sample2"
one_group = paste0(sample_name,"_Xenium_v2")
```

```{r outputPrefix}
my_output_prefix_figure = paste0(my_output_prefix,"figures/",one_group,"/")
if(!dir.exists(my_output_prefix_figure)){dir.create(my_output_prefix_figure,recursive = TRUE)}
```


Load the Xenium data
```{r ReadInData}
raw_data_dir = "SpatialData/PublicData10x/data/"
path <- paste0(raw_data_dir,"/","Sample2/outs")
```

```{r ReadInData, eval=eval_read_in_raw_data}
xenium.obj.raw <- MyLoadXenium(path)
```


Number of genes and cells before filtering
```{r}
dim(xenium.obj.raw)
```


<!-- Supervised labeling & label transfer -->
read in cell type annotation from 10x
```{r}
filename_csv = paste0(data_dir,"/","Cell_type_annotation/",
                      "Cell_type_annotation_Xenium_Sample2_unsupervised.csv")
cell_type_anno_10x = read.csv(filename_csv)
cell_type_anno_10x$Barcode = cell_type_anno_10x$Barcode
rownames(cell_type_anno_10x) = as.character(cell_type_anno_10x$Barcode)
dim(cell_type_anno_10x)
```


raw cell type annotation (not recomment to use the unspervised annotation)
```{r}
table(cell_type_anno_10x$Cluster)
```

<!-- check cell names: should be true-->
```{r,include=FALSE}
all.equal(colnames(xenium.obj.raw),rownames(cell_type_anno_10x))
```


Xenium object for analysis
```{r}
xenium.obj = xenium.obj.raw
```

Add sample name information
```{r}
xenium.obj@meta.data$sample_name = sample_name
```


Remove cells annotated as "Unlabeled"
```{r RemoveZeroCounts}
cells_kept = rownames(subset(cell_type_anno_10x, !(Cluster %in% "Not Plotted") ))
xenium.obj <- xenium.obj[,cells_kept]
```


Number of genes and cells after filtering
```{r}
dim(xenium.obj)
```


<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

### Analysis of Xenium data

Set parameters
```{r}
n_dims = 30
n_resolution = 0.3
```

Seurat pipeline for data analysis
```{r SeuratPipeline, results='hide',eval=eval_Seurat}
xenium.obj <- NormalizeData(xenium.obj, assay = "Xenium")
xenium.obj <- ScaleData(xenium.obj, assay = "Xenium")
xenium.obj <- RunPCA(xenium.obj, npcs = n_dims, features = rownames(xenium.obj))
xenium.obj <- RunUMAP(xenium.obj, dims = 1:n_dims)
xenium.obj <- FindNeighbors(xenium.obj, reduction = "pca", dims = 1:n_dims)
xenium.obj <- FindClusters(xenium.obj, resolution = n_resolution)
```


Backup raw clusters
```{r}
xenium.obj$seurat_clusters.raw <- xenium.obj$seurat_clusters
```


Find sub clusters for cluster 1. Decide to run FindSubCluster after looking the downstream feature plots of marker genes.
```{r}
xenium.obj =  FindSubCluster(xenium.obj,
                              graph.name = "Xenium_snn",
                              cluster = "1",
                              resolution = 0.1)
```

Assign sub clusters to seurat clusters and active.ident
```{r}
ordered_numer = sort(table(xenium.obj$sub.cluster),decreasing = TRUE)
xenium.obj$sub.cluster = factor(xenium.obj$sub.cluster,
                                 levels = names(ordered_numer) )

xenium.obj$seurat_clusters = factor(xenium.obj$sub.cluster,
                                     levels = names(ordered_numer),
                                     labels = 0:(length(ordered_numer)-1) )
xenium.obj@active.ident <- xenium.obj$seurat_clusters
```


Number of cells in each cluster
```{r}
table(xenium.obj$seurat_clusters)
```


<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


### Cell type annotation

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### Manual using known marker genes

Method for cell type prediction
```{r}
pred_method = "Manual"
```

Manual assign cell types
```{r}
xenium.obj$predicted_cell_type_Manual = xenium.obj$seurat_clusters
levels(xenium.obj$predicted_cell_type_Manual) =  c("T",
                                                   "Fibroblast",
                                                   "Macrophage",
                                                   "Endothelial",
                                                   "Tumor",
                                                   "Basal",
                                                   "LP",
                                                   "T",  # ? (? means hard to decide the cell type)
                                                   "ML",
                                                   "T", # ?
                                                   "B",
                                                   "Plasma",
                                                   "Mast",
                                                   "Tumor",
                                                   "Adipocyte",
                                                   "T")
nlevels(xenium.obj$predicted_cell_type_Manual)
```


Order cell type by cell number
```{r}
ordered_numer = sort(table(xenium.obj$predicted_cell_type_Manual),decreasing = TRUE)
xenium.obj$predicted_cell_type_Manual = factor(xenium.obj$predicted_cell_type_Manual, 
                                               levels = names(ordered_numer) )
table(xenium.obj$predicted_cell_type_Manual)
```


DimPlot by predicted cell types
```{r DimPlotPredictedCellType, results='hide',include=FALSE}
p <- list()
my_group_by <- c("predicted_cell_type_Manual")
my_colors  <- my_colors_20

for (i in 1:length(my_group_by)) {
  p[[i]] <- DimPlot(xenium.obj,
                    pt.size=point_size,
                    reduction="umap",
                    group.by=my_group_by[i],
                    shuffle=TRUE,
                    cols=my_colors) +
    my_theme_ggplot +
    ggplot2::labs(title=pred_method)
}

my_output_filename = paste0(my_output_prefix_figure,
                            one_group,"_",
                            "UMAP_plot_predicted_cell_type_",pred_method,
                            ".png")

png(my_output_filename,
    width = 4.2*1*1.5, height = 3*1*1.5, units = "in",res = 300)
patchwork::wrap_plots(p, nrow=1, byrow=FALSE)
dev.off()
```

```{r, echo = FALSE, out.width='80%', fig.align='center'}
knitr::include_graphics(my_output_filename)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

Create a list to save running time
```{r}
running_time = list()
```

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### SingleR

Method for cell type prediction
```{r}
pred_method = "SingleR"
```

read in scRNASeq reference data
```{r}
filename_rds = paste0(rds_res_dir,"/", 
                      "data_seurat_cell_type_Flex_Sample1",".rds")
data_seurat = readRDS(file = filename_rds)
```

```{r,include=FALSE}
filename_rds = ""
```


Number of singlets and doublets in the ref single cell data
```{r}
table(data_seurat$db_type)
```


convert seurat object into sce object
```{r sceSinglet}
sce_ref = data_seurat[,data_seurat$db_type=="singlet"]
sce_ref <- as.SingleCellExperiment(DietSeurat(sce_ref,graphs=c("pca","umap")) )
```


sce reference data for cell type annotation
```{r sceRef}
common_genes = intersect( rownames(xenium.obj), rownames(sce_ref))
sce_ref = sce_ref[common_genes,]
```

xenium object to sce object
```{r sceTest}
sce_test  = as.SingleCellExperiment(DietSeurat(xenium.obj,graphs=c("pca","umap")) )
```

predict cell type using singleR
```{r PredictCellType, eval=eval_predict_cell_type_SingleR}
t_start <- Sys.time()

pred_celltype <- SingleR(test=sce_test, ref=sce_ref, 
                         labels=sce_ref$cell_type_label, de.method="wilcox")

t_end <- Sys.time()
t_diff <- difftime(t_end, t_start,units = "mins")
running_time[[pred_method]][["t_start"]]  <- t_start
running_time[[pred_method]][["t_end"]]  <- t_end
running_time[[pred_method]][["t_diff"]]  <- t_diff
```

<!-- save pred_celltype into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
saveRDS(pred_celltype, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
pred_celltype = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
saveRDS(running_time[[pred_method]], file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
running_time[[pred_method]] = readRDS(file = filename_rds)
```


Add predicted cell type to xenium object
```{r AddPredictedCellType}
xenium.obj@meta.data$predicted_cell_type_SingleR = factor(pred_celltype$labels)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### Azimuth

Method for cell type prediction
```{r}
pred_method = "Azimuth"
```

azimuth reference dir
```{r}
ref_azimuth_dir = paste0(my_output_prefix,"ref_azimuth/Flex_Sample1")
if(!dir.exists(ref_azimuth_dir)){dir.create(ref_azimuth_dir,recursive = TRUE)}
```


Predict cell types using azimuth
```{r,results='hide',eval=eval_predict_cell_type_azimuth}
t_start <- Sys.time()

b = RunAzimuth(xenium.obj,reference = ref_azimuth_dir,assay = "Xenium")

t_end <- Sys.time()
t_diff <- difftime(t_end, t_start,units = "mins")
running_time[[pred_method]][["t_start"]]  <- t_start
running_time[[pred_method]][["t_end"]]  <- t_end
running_time[[pred_method]][["t_diff"]]  <- t_diff

pred_celltype_azimuth = b@meta.data[,c("predicted.cell_type_label.score",
                                       "predicted.cell_type_label",
                                       "mapping.score")]
```


<!-- save pred_celltype into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
saveRDS(pred_celltype_azimuth, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
pred_celltype_azimuth = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
saveRDS(running_time[[pred_method]], file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
running_time[[pred_method]] = readRDS(file = filename_rds)
```

assign predicted cell type to xenium object
```{r}
xenium.obj$predicted_cell_type_Azimuth <- pred_celltype_azimuth$predicted.cell_type_label
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### scmap

Method for cell type prediction
```{r}
pred_method = "scmapCell"
```

convert seurat object into sce object
```{r}
sce = data_seurat[,data_seurat$db_type=="singlet"]
sce <- as.SingleCellExperiment(DietSeurat(sce,graphs=c("pca","umap")) )
```

sce reference data for cell type annotation
```{r}
common_genes = intersect( rownames(xenium.obj), rownames(sce))
sce_ref = sce[common_genes,]
rowData(sce_ref)$feature_symbol <- rownames(sce_ref)
```

xenium object to sce object
```{r}
sce_query  = as.SingleCellExperiment(DietSeurat(xenium.obj,graphs=c("pca","umap")) )
rowData(sce_query)$feature_symbol <- rownames(sce_query)
```

select the most informative features
```{r}
sce_ref <- selectFeatures(sce_ref, suppress_plot = TRUE)
```

set a seed
```{r}
set.seed(42)
```

create scmap-cell index
```{r}
sce_ref <- indexCell(sce_ref)
```

```{r}
names(metadata(sce_ref)$scmap_cell_index)
```


Projection to indexes
```{r, eval=eval_predict_cell_type_scmapCell}
t_start <- Sys.time()

scmapCell_results <- scmapCell(
  projection = sce_query, 
  index_list = list(
    ref_sc = metadata(sce_ref)$scmap_cell_index
  )
)

t_end <- Sys.time()
t_diff <- difftime(t_end, t_start,units = "mins")
running_time[[pred_method]][["t_start"]]  <- t_start
running_time[[pred_method]][["t_end"]]  <- t_end
running_time[[pred_method]][["t_diff"]]  <- t_diff
```


<!-- save scmapCell_results into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_scmapCell}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
saveRDS(scmapCell_results, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_scmapCell}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
scmapCell_results = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_scmapCell}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
saveRDS(running_time[[pred_method]], file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_scmapCell}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
running_time[[pred_method]] = readRDS(file = filename_rds)
```


Cluster annotation
```{r}
scmapCell_clusters <- scmapCell2Cluster(
  scmapCell_results, 
  list(
    as.character(colData(sce_ref)$cell_type_label)
  )
)
```


assign predicted cell type to xenium object
```{r}
xenium.obj$predicted_cell_type_scmapCell <- scmapCell_clusters$scmap_cluster_labs
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### RCTD

Method for cell type prediction
```{r}
pred_method = "RCTD"
```

Query counts for RCTD
```{r}
library(spacexr)
query.counts <- GetAssayData(xenium.obj, assay = "Xenium", slot = "counts")
coords <- GetTissueCoordinates(xenium.obj, which = "centroids")
rownames(coords) <- coords$cell
coords$cell <- NULL
query <- SpatialRNA(coords, query.counts, colSums(query.counts))
```

RCTD reference dir
```{r outputPrefix}
ref_RCTD_dir = paste0(my_output_prefix,"ref_RCTD/Flex_Sample1")
if(!dir.exists(ref_RCTD_dir)){dir.create(ref_RCTD_dir,recursive = TRUE)}
```

Reference for RCTD
```{r}
ref_RCTD = readRDS(file = paste0(ref_RCTD_dir,"/","reference.rds"))
```

library size range of spatial data
```{r}
range(query@nUMI)
```

run RCTD with many cores
```{r,results='hide'}
RCTD <- create.RCTD(query, ref_RCTD, 
                    max_cores = 8,
                    UMI_min = 0,     # min UMI to filter cells in spatial data
                    counts_MIN = 0,  # minimum total counts per pixel of genes
                    gene_cutoff = 0, # minimum normalized gene expression for genes to be included
                    fc_cutoff = 0,
                    fc_cutoff_reg = 0,
                    UMI_min_sigma = 1,
                    CELL_MIN_INSTANCE=10)
```

```{r}
dim(query@counts)
```

```{r}
dim(RCTD@spatialRNA@counts)
```

```{r,results='hide', eval=eval_predict_cell_type_RCTD}
t_start <- Sys.time()

RCTD <- run.RCTD(RCTD, doublet_mode = "doublet")

t_end <- Sys.time()
t_diff <- difftime(t_end, t_start,units = "mins")
running_time[[pred_method]][["t_start"]]  <- t_start
running_time[[pred_method]][["t_end"]]  <- t_end
running_time[[pred_method]][["t_diff"]]  <- t_diff
```

<!-- save into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
saveRDS(RCTD, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
RCTD = readRDS(file = filename_rds)
```

<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
saveRDS(running_time[[pred_method]], file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
running_time[[pred_method]] = readRDS(file = filename_rds)
```

assign predicted cell type to xenium object
```{r}
annotations.df <- RCTD@results$results_df
annotations <- annotations.df$first_type
names(annotations) <- rownames(annotations.df)
all.equal(names(annotations),colnames(xenium.obj))
xenium.obj$predicted_cell_type_RCTD <- annotations
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->



<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### scPred

Method for cell type prediction
```{r}
pred_method = "scPred"
```


240318-notes: scPred does not require common genes between reference and query
Reference scRNASeq dataset
```{r}
common_genes = intersect( rownames(xenium.obj), rownames(data_seurat))
reference = data_seurat[common_genes,]
```


Let feature space to train the classifiers. By default, scPred will use all principal components.
```{r,results='hide'}
library(scPred)
reference <- getFeatureSpace(reference, "cell_type_label")
```


Query dataset
```{r}
query = xenium.obj[common_genes,]
```

Train a model and predict cell types
```{r,eval=eval_predict_cell_type_scPred}
t_start <- Sys.time()
# Train the classifiers, By default, scPred will use a support vector machine with a radial kernel.
reference <- trainModel(reference)

# predict cell types
query <- scPredict(query, reference)

t_end <- Sys.time()
t_diff <- difftime(t_end, t_start,units = "mins")
running_time[[pred_method]][["t_start"]]  <- t_start
running_time[[pred_method]][["t_end"]]  <- t_end
running_time[[pred_method]][["t_diff"]]  <- t_diff
```


<!-- save reference into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_scPred}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","reference_trained_",pred_method,".rds")
saveRDS(reference, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_scPred}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","reference_trained_",pred_method,".rds")
reference = readRDS(file = filename_rds)
```


<!-- save into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_scPred}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
saveRDS(query, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_scPred}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","pred_celltype_",pred_method,".rds")
query = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_scPred}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
saveRDS(running_time[[pred_method]], file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_scPred}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_",pred_method,".rds")
running_time[[pred_method]] = readRDS(file = filename_rds)
```

assign predicted cell type to xenium object
```{r}
xenium.obj$predicted_cell_type_scPred <- query$scpred_prediction
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- Save running time -->
```{r,include=FALSE}
filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_","combined",".rds")
saveRDS(running_time, file = filename_rds)
```

<!-- save Xenium data into rds -->
```{r, include=FALSE,eval=eval_save_xenium_obj}
filename_rds = paste0(rds_res_dir,"/", "xenium_obj_cell_type","_", one_group,".rds")
saveRDS(xenium.obj, file = filename_rds)
```


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

<!-- settings for plotting -->

<!-- Set figure with and heigt for image dim plot -->
```{r,include=FALSE}
do_flip = TRUE
if(do_flip){
  fig_w = 3*2*1.6
  fig_h = 3*1.6
}else{
  fig_w = 3*2*1
  fig_h = 3*2*1
}
```


<!-- Set point size -->
```{r,include=FALSE}
point_size = 1.5
```
<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

<!-- figures for 2025. BMC Bioinformatics. paper. -->

## Figures

### Figure 7

Methods for cell type annotation
```{r}
pred_methods = c("Manual","SingleR","Azimuth","RCTD","scPred","scmapCell")
```

Set colors
```{r}
all_cell_types = list()
for (pred_method in pred_methods) {
  all_cell_types[[pred_method]] = 
    levels(factor(unlist(xenium.obj[[paste0("predicted_cell_type_",
                                                  pred_method)]])))
}
all_cell_types = unique(do.call(c,all_cell_types)) # use order of first method (Manual)
# all_cell_types = sort( all_cell_types ) # use alphabetic order
all_cell_types = all_cell_types[all_cell_types != "unassigned"]
all_cell_types = c(all_cell_types,"unassigned") # put "unassigned" last
all_cell_types 
```


```{r}
my_named_colors = my_colors_20[1:length(all_cell_types)]
names(my_named_colors) = all_cell_types
my_named_colors["unassigned"] <- "black"
```

Image DimPlot of predicted cell type
```{r,include=FALSE}
p=list()

for (pred_method in pred_methods) {
  xenium.obj$predicted_cell_type = unlist(xenium.obj[[paste0("predicted_cell_type_",
                                                      pred_method)]])
  # order cell type by alphabetic order
  # xenium.obj$predicted_cell_type = factor(as.character(xenium.obj$predicted_cell_type))
  
  # (Or) order cell type by cell number
  ordered_numer = sort(table(xenium.obj$predicted_cell_type),decreasing = TRUE)
  xenium.obj$predicted_cell_type = factor(xenium.obj$predicted_cell_type,
                                          levels = names(ordered_numer) )
  
  my_colors  <- my_named_colors[levels(xenium.obj$predicted_cell_type)]
  
  
  p[[pred_method]] <- ImageDimPlot(xenium.obj,
                    group.by = "predicted_cell_type",
                    cols = my_colors, 
                    axes = TRUE) + 
      theme_bw() + my_theme_ggplot + 
    labs(title = pred_method,fill="")  + 
    theme(plot.title=element_text(size=25, face="bold",hjust=0.5)) + 
    
    theme(legend.title=element_text(size=12*0.8, face="bold"),  # modify legend size for thesis
          legend.text=element_text(size=12*0.8),
          legend.key.size=unit(1*0.8, "cm"),
          legend.key.width=unit(0.45, "cm"),
          legend.key.height=unit(0.45, "cm"),
          legend.spacing.y=unit(0.45, "cm"),
          legend.spacing.x=unit(0.45, "cm")
          )
  
  
  if(do_flip){p[[pred_method]] <- p[[pred_method]] + coord_flip()}


}


n_col = 2
n_row = ceiling(length(pred_methods)/n_col)
scale_file = 0.8

my_output_filename = paste0(my_output_prefix_figure,one_group,"_",
                            "ImageDimPlot_of_predicted_cell_type",".png")
png(my_output_filename, width = fig_w*n_col*scale_file,
    height = fig_h*n_row*scale_file, units = "in",res = 300)
print(patchwork::wrap_plots(p, ncol=n_col, byrow=TRUE))
dev.off()
```


```{r, echo = FALSE, out.width='60%', fig.align='center'}
my_output_filename = paste0(my_output_prefix_figure,one_group,"_",
                            "ImageDimPlot_of_predicted_cell_type",".png")
knitr::include_graphics(my_output_filename)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

### Figure 8

#### 8A

Number of cells of each predicted cell type
```{r}
tab = list()
pred_methods = c("Manual","SingleR","Azimuth","RCTD","scPred","scmapCell")

for (i in pred_methods) {
  tab[[i]] = as.data.frame( table(xenium.obj[[paste0("predicted_cell_type_",i)]]) )
  tab[[i]]$Method = i
  colnames(tab[[i]])[1] <- "predicted_cell_type"
}

tab = do.call(rbind.data.frame,tab)

# make sure the levels order is the same with pred_methods!
tab$Method = factor(tab$Method,levels = pred_methods)
```


Percentage of each cell type in by each method 
```{r}
ratio <- tapply(tab$Freq, tab$Method, 
                function(x){round(100*x/sum(x),2)})
ratio <- do.call(c,ratio)
tab$Ratio <- ratio
```

relevel predicted cell type
```{r}
tab$predicted_cell_type <- factor(tab$predicted_cell_type,levels = all_cell_types)
```

Barplot of cell number of predicted cell types
```{r}
my_colors = my_named_colors 
p1 <- ggplot(data=tab, aes(x=Method, y=Freq, fill=predicted_cell_type)) + 
  scale_fill_manual(values=my_colors) + geom_bar(stat="identity") + 
  labs(x=NULL, y="Number", fill="Predicted cell type") + 
  my_theme_ggplot + RotatedAxis() + 
  theme(legend.title=element_text(size=12*0.8, face="bold"),
        legend.text=element_text(size=12*0.8),
        legend.key.size=unit(1*0.8, "cm"),
        legend.key.width=unit(0.45, "cm"),
        legend.key.height=unit(0.45, "cm"),
        legend.spacing.y=unit(0.45, "cm"),
        legend.spacing.x=unit(0.45, "cm")
        )
```

#### 8B

```{r}
list_similarity = list()

ref_based_methods = c("SingleR","Azimuth","RCTD","scPred","scmapCell")

a = as.character(xenium.obj$predicted_cell_type_Manual)

for (one_method in ref_based_methods ) {
  b = as.character(xenium.obj@meta.data[,paste0("predicted_cell_type_",one_method)])
  
  tab_num <- table(a == b)
  
  list_similarity[[one_method]] <- tab_num
}

# use rbind
tab_similarity = do.call(rbind,list_similarity)
tab_similarity
```

Cell proportion of each cell type in each sample
```{r}
tab_ratio <- round(100*tab_similarity/rowSums(tab_similarity), 2)
tab_ratio <- as.data.frame.matrix(tab_ratio)
tab_ratio
```

```{r}
tab_ratio <- cbind(tab_ratio, Method=rownames(tab_ratio))
tab_ratio$Method <- factor(tab_ratio$Method, 
                           levels= ref_based_methods)
tab_ratio = tab_ratio[,c("TRUE","Method")]
colnames(tab_ratio)[1] <- "Similarity"
tab_ratio
```


```{r}
p2 <- ggplot(data=tab_ratio, aes(x=Method, y=Similarity, fill=Method)) + 
  scale_fill_manual(values=my_colors_20) + geom_bar(stat="identity") + 
  geom_text(aes(label = Similarity), vjust=-0.5, size=5) + # add this line after geom_bar to avoid incorrect visualization
  ylim(c(0,80)) + 
  labs(y="Percentage (%)", x=NULL, fill="Method", title = "Similarity with Manual") + 
  my_theme_ggplot + RotatedAxis() # + NoLegend()

```


#### 8C

<!-- read in running time -->
<!-- ```{r,include=FALSE} -->
<!-- filename_rds = paste0(rds_res_dir,"/", one_group,"_","running_time_","combined",".rds") -->
<!-- running_time = readRDS(filename_rds) -->
<!-- ``` -->

print running time
```{r}
time_df = do.call(rbind,lapply(running_time,as.data.frame))
a = round(as.numeric(time_df$t_diff),2)
a = cbind.data.frame("Method" = rownames(time_df),
                     "RunningTime" = a )
a$Method = factor(a$Method,levels = c("SingleR","Azimuth",
                                      "RCTD","scPred","scmapCell"))
a 
```

Barplot of running time of each method
```{r, include=FALSE}
p3 <- ggplot(data=a, aes(x=Method, y=RunningTime,fill=Method)) + 
  scale_fill_manual(values=my_colors_20) + geom_bar(stat="identity") + 
  labs(x=NULL, y="Running Time (Minutes)", fill="") + 
  my_theme_ggplot + RotatedAxis()

```


Combine p1, p2 and p3
```{r}
n_row = 2
n_col = 1
scale_file = 1.5

n_fontsize = 20

my_plot = ggpubr::ggarrange(
  p1, p2, p3,
  # nrow = n_row,
  labels = c("A","B","C"), # Labels 
  font.label = list(size = n_fontsize, face = "bold", color ="black")
) 


my_output_filename = paste0(my_output_prefix_figure,one_group,"_",
                            "Figure7",".png")
png(my_output_filename, width = 8*n_col*scale_file,
    height = 3*n_row*scale_file, units = "in",res = 300)
my_plot
dev.off()
```

```{r, echo = FALSE, out.width='60%', fig.align='center'}
my_output_filename = paste0(my_output_prefix_figure,one_group,"_",
                            "Figure7",".png")
knitr::include_graphics(my_output_filename)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

\pagebreak

