
<!-- Author: Jinming Cheng -->
<!-- Date: 17 July 2025 -->

<!-- Notes: only SingleR, Azimuth and RCTD were selected for running multiple times  -->
<!-- using subsetted references -->

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

<!-- # Sample1 Rep1 (v2) -->

<!-- ## 10x Xenium -->

### Multiple runs

Run cell type annotation multiple times for SingleR, Azimuth and RCTD.


```{r,include=FALSE}
eval_sample_cells <- FALSE
eval_predict_cell_type_SingleR = FALSE
eval_predict_cell_type_azimuth = TRUE
eval_predict_cell_type_RCTD = FALSE
eval_save_rdata = TRUE
```


```{r,include=FALSE}
load("results/00_prepare.rdata")
```

Load packages
```{r}
library(Seurat)
library(ggplot2)
library(edgeR)
library(SingleR)
library(Azimuth)
library(scmap)
library(spacexr)
library(scPred)
```


<!-- Prefix of output file names -->
```{r,include=FALSE}
sample_name = "Rep1"
one_group = paste0(sample_name,"_Xenium_v2")
```

```{r,include=FALSE}
my_output_prefix_figure = paste0(my_output_prefix,"figures/",one_group,"/")
if(!dir.exists(my_output_prefix_figure)){dir.create(my_output_prefix_figure,recursive = TRUE)}
```

<!-- Read in final xenium object -->
```{r,include=FALSE}
filename_rds = paste0(rds_res_dir,"/", "xenium_obj_cell_type","_", one_group,".rds")
xenium.obj = readRDS( filename_rds)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->



<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

<!-- read in scRNASeq reference data -->
```{r,include=FALSE}
filename_rds = paste0(rds_res_dir,"/", 
                      "data_seurat_cell_type_Flex_Sample1",".rds")
data_seurat = readRDS(file = filename_rds)
filename_rds = ""
```

<!-- Keep singlets only -->
```{r,include=FALSE}
data_seurat <- data_seurat[,data_seurat$db_type=="singlet"]
```

```{r}
dim(data_seurat)
```


```{r}
table(data_seurat$cell_type_label)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### Sample cells

Function definition to sample cells for multiple groups ensuring a minimum number of cells in each group

```{r}
# Load necessary library
library(dplyr)

# Example data frame creation (replace this with your actual data)
# set.seed(42) # for reproducibility
# data <- data.frame(
#   cell_id = 1:10000,
#   cell_type = sample(c("Type1", "Type2", "Type3", "Type4"), 10000, replace = TRUE)
# )

# Function to sample cells
sample_cells <- function(data, 
                         n_total_cells = 5000, 
                         min_per_group = 50,
                         seed = 42) {
  
  # for reproducibility
  set.seed(seed) # for reproducibility
  
  # Ensuring each group has at least min_per_group cells
  initial_cells <- data %>%
    dplyr::group_by(cell_type) %>%
    dplyr::slice_sample(n = min_per_group)

  # Additional sampling if n_total_cells > number of cells already sampled
  n_additional_needed <- n_total_cells - nrow(initial_cells)
  if (n_additional_needed > 0) {
    # Exclude already selected cells
    remaining_data <- dplyr::anti_join(data, initial_cells, by = "cell_id")

    # Sample additional cells from the remaining data
    additional_cells <- remaining_data %>%
      dplyr::sample_n(n_additional_needed)
    
    # Combine initial and additional samples
    final_cells <- dplyr::bind_rows(initial_cells, additional_cells)
  } else {
    final_cells <- initial_cells
  }
  
  return(final_cells)
}

# # Use the function to sample cells
# selected_cells <- sample_cells(data)
# 
# # View the results
# table(selected_cells$cell_type)
```


Data used for sampling
```{r}
data <- data.frame(
  cell_id = colnames(data_seurat),
  cell_type = data_seurat$cell_type_label
)
```


Number of cells for sampleing
```{r}
n_cells_sampling = seq(4000,20000,length.out =5)
n_cells_sampling
```

Sample cells
```{r,eval=eval_sample_cells}
list_selected_cells <- list()

for (k in  n_cells_sampling) {
  # Use the function to sample cells
  selected_cells <- sample_cells(data,
                                 n_total_cells = k,
                                 min_per_group = 50,
                                 seed = 42)
  list_selected_cells[[paste0(k,"cells")]] <- selected_cells
}
```

<!-- save list_selected_cells into rds -->
```{r,include=FALSE,eval=eval_sample_cells}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "list_selected_cells_from_sampling",".rds")
saveRDS(list_selected_cells, file = filename_rds)
```


<!-- read list_selected_cells into rds -->
```{r,include=FALSE,eval=!eval_sample_cells}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "list_selected_cells_from_sampling",".rds")
list_selected_cells <- readRDS(file = filename_rds)
```


Cell number of each group from 1st sampling
```{r}
table(list_selected_cells[[1]]$cell_type)
```

```{r}
names(list_selected_cells)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


Create a list to save time
```{r}
multiple_running_time = list()
```


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### SingleR

Method for cell type prediction
```{r}
pred_method = "SingleR"
```


Predict cell type
```{r,eval=eval_predict_cell_type_SingleR}
multiple_pred_celltype = list()

# test dataset
sce_test  = as.SingleCellExperiment(DietSeurat(xenium.obj,graphs=c("pca","umap")) )


for (i in names(list_selected_cells) ) {
  
  # keep a subset of cells
  cells_kept = list_selected_cells[[i]]$cell_id
  data_seurat_sub = data_seurat[,cells_kept]
  
  # sce reference
  sce_ref = data_seurat_sub
  sce_ref <- as.SingleCellExperiment(DietSeurat(sce_ref,graphs=c("pca","umap")) )
  common_genes = intersect( rownames(xenium.obj), rownames(sce_ref))
  sce_ref = sce_ref[common_genes,]
  
  
  # run Single R and record running time
  t_start <- Sys.time()
  multiple_pred_celltype[[i]] <- SingleR(test=sce_test, ref=sce_ref,
                                         labels=sce_ref$cell_type_label, 
                                         de.method="wilcox")
  t_end <- Sys.time()
  t_diff <- difftime(t_end, t_start,units = "mins")
  multiple_running_time[[pred_method]][[i]][["t_start"]]  <- t_start
  multiple_running_time[[pred_method]][[i]][["t_end"]]  <- t_end
  multiple_running_time[[pred_method]][[i]][["t_diff"]]  <- t_diff
}

multiple_pred_celltype_SingleR <- multiple_pred_celltype
```


<!-- save pred_celltype into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_pred_celltype_",pred_method,".rds")
saveRDS(multiple_pred_celltype_SingleR, file = filename_rds)
```

<!-- read in rds -->
```{r,include=FALSE,eval=!eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_pred_celltype_",pred_method,".rds")
multiple_pred_celltype_SingleR = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_",pred_method,".rds")
saveRDS(multiple_running_time[[pred_method]], file = filename_rds)
```

<!-- read in rds -->
```{r,include=FALSE,eval=!eval_predict_cell_type_SingleR}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_",pred_method,".rds")
multiple_running_time[[pred_method]] = readRDS(file = filename_rds)
```


Add predicted cell type to xenium object
```{r}
for (i in names(list_selected_cells) ) {
  one_col_name = paste0("predicted_cell_type_SingleR","_",i)
  xenium.obj@meta.data[[one_col_name]] = 
    multiple_pred_celltype_SingleR[[i]]$labels
}
```


```{r,include=FALSE,eval=FALSE}
table(xenium.obj@meta.data[["predicted_cell_type_SingleR_4000cells"]]==xenium.obj@meta.data[["predicted_cell_type_SingleR_8000cells"]])
```


print running time (minutes)
```{r}
time_df = do.call(rbind,lapply(multiple_running_time[[pred_method]],as.data.frame))
a = round(as.numeric(time_df$t_diff),2)
a = cbind.data.frame("NCells" = rownames(time_df),
                     "RunningTime" = a )
a 
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### Azimuth

Method for cell type prediction
```{r}
pred_method = "Azimuth"
```


Predict cell types using azimuth
```{r,results='hide',eval=eval_predict_cell_type_azimuth}

multiple_pred_celltype = list()

for (i in names(list_selected_cells) ) {
  # azimuth reference dir
  ref_azimuth_dir = paste0(my_output_prefix,"ref_azimuth/Flex_Sample1","_",i)
  
  # record running time
  t_start <- Sys.time()
  b = RunAzimuth(xenium.obj,reference = ref_azimuth_dir,assay = "Xenium")
  t_end <- Sys.time()
  t_diff <- difftime(t_end, t_start,units = "mins")
  multiple_running_time[[pred_method]][[i]][["t_start"]]  <- t_start
  multiple_running_time[[pred_method]][[i]][["t_end"]]  <- t_end
  multiple_running_time[[pred_method]][[i]][["t_diff"]]  <- t_diff
  
  multiple_pred_celltype[[i]] = b@meta.data[,c("predicted.cell_type_label.score",
                                               "predicted.cell_type_label",
                                               "mapping.score")]
}

multiple_pred_celltype_azimuth <- multiple_pred_celltype

```


<!-- save pred_celltype into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_pred_celltype_",pred_method,".rds")
saveRDS(multiple_pred_celltype_azimuth, file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_pred_celltype_",pred_method,".rds")
multiple_pred_celltype_azimuth = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_",pred_method,".rds")
saveRDS(multiple_running_time[[pred_method]], file = filename_rds)
```

```{r,include=FALSE,eval=!eval_predict_cell_type_azimuth}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_",pred_method,".rds")
multiple_running_time[[pred_method]] = readRDS(file = filename_rds)
```


Add predicted cell type to xenium object
```{r}
for (i in names(list_selected_cells) ) {
  one_col_name = paste0("predicted_cell_type_Azimuth","_",i)
  xenium.obj@meta.data[[one_col_name]] =
    multiple_pred_celltype_azimuth[[i]]$predicted.cell_type_label
}
```


print running time (minutes)
```{r}
time_df = do.call(rbind,lapply(multiple_running_time[[pred_method]],as.data.frame))
a = round(as.numeric(time_df$t_diff),2)
a = cbind.data.frame("NCells" = rownames(time_df),
                     "RunningTime" = a )
a 
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

#### RCTD

Method for cell type prediction
```{r}
pred_method = "RCTD"
```


Predict cell types
```{r,results='hide', eval=eval_predict_cell_type_RCTD}
multiple_pred_celltype = list()

# Query counts for RCTD
query.counts <- GetAssayData(xenium.obj, assay = "Xenium", slot = "counts")
coords <- GetTissueCoordinates(xenium.obj, which = "centroids")
rownames(coords) <- coords$cell
coords$cell <- NULL
query <- SpatialRNA(coords, query.counts, colSums(query.counts))

for (i in names(list_selected_cells) ) {
  
  # RCTD reference dir
  ref_RCTD_dir = paste0(my_output_prefix,"ref_RCTD/Flex_Sample1","_",i)
  # RCTD reference
  ref_RCTD = readRDS(file = paste0(ref_RCTD_dir,"/","reference.rds"))
  
  # create RCTD object
  RCTD <- create.RCTD(query, ref_RCTD, 
                      max_cores = 8,
                      UMI_min = 0,     # min UMI to filter cells in spatial data
                      counts_MIN = 0,  # minimum total counts per pixel of genes
                      gene_cutoff = 0, # minimum normalized gene expression for genes to be included
                      fc_cutoff = 0,
                      fc_cutoff_reg = 0,
                      UMI_min_sigma = 1,
                      CELL_MIN_INSTANCE=10)
  
  # record running time
  t_start <- Sys.time()
  multiple_pred_celltype[[i]] <- run.RCTD(RCTD, doublet_mode = "doublet")
  t_end <- Sys.time()
  t_diff <- difftime(t_end, t_start,units = "mins")
  multiple_running_time[[pred_method]][[i]][["t_start"]]  <- t_start
  multiple_running_time[[pred_method]][[i]][["t_end"]]  <- t_end
  multiple_running_time[[pred_method]][[i]][["t_diff"]]  <- t_diff
  
}

multiple_pred_celltype_RCTD <- multiple_pred_celltype
```


<!-- save into rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_pred_celltype_",pred_method,".rds")
saveRDS(multiple_pred_celltype_RCTD, file = filename_rds)
```

<!-- read in rds -->
```{r,include=FALSE,eval=!eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_pred_celltype_",pred_method,".rds")
multiple_pred_celltype_RCTD = readRDS(file = filename_rds)
```


<!-- save running time rds -->
```{r,include=FALSE,eval=eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_",pred_method,".rds")
saveRDS(multiple_running_time[[pred_method]], file = filename_rds)
```

<!-- read in rds -->
```{r,include=FALSE,eval=!eval_predict_cell_type_RCTD}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_",pred_method,".rds")
multiple_running_time[[pred_method]] = readRDS(file = filename_rds)
```


Add predicted cell type to xenium object
```{r}

for (i in names(list_selected_cells) ) {
  one_col_name = paste0("predicted_cell_type_RCTD","_",i)
  annotations.df <- multiple_pred_celltype_RCTD[[i]]@results$results_df
  annotations <- annotations.df$first_type
  names(annotations) <- rownames(annotations.df)
  # all.equal(names(annotations),colnames(xenium.obj))
  xenium.obj@meta.data[[one_col_name]] <- annotations

}

```


print running time (minutes)
```{r}
time_df = do.call(rbind,lapply(multiple_running_time[[pred_method]],as.data.frame))
a = round(as.numeric(time_df$t_diff),2)
a = cbind.data.frame("NCells" = rownames(time_df),
                     "RunningTime" = a )
a 
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->


<!-- Save running time -->
```{r,include=FALSE}
filename_rds = paste0(rds_res_dir,"/", one_group,"_",
                      "multiple_runs_running_time_","combined",".rds")
saveRDS(multiple_running_time, file = filename_rds)
```


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- %                                                 % -->

Comparing runtime of different methods (multipe runs)

#### Figure 6B

print running time (minutes)
```{r}
df_time = list()

for (i in names(multiple_running_time)) {
  time_df = do.call(rbind,lapply(multiple_running_time[[i]],as.data.frame))
  a = round(as.numeric(time_df$t_diff),2)
  a = cbind.data.frame("NCells" = rownames(time_df),
                     "RunningTime" = a )
  a$Method = i 
  df_time[[i]] <- a
}

df_time = do.call(rbind,df_time)
rownames(df_time) <- NULL
df_time$NCells <- as.numeric(gsub("cells","",df_time$NCells))
df_time$Method <- factor(df_time$Method, levels = names(multiple_running_time))
df_time
```


linechart of multiple running time
```{r, include=FALSE}
p <- ggplot(data=df_time, aes(x=NCells, y=RunningTime,color=Method)) + 
  scale_color_manual(values=my_colors_20) + 
  geom_point(size=2) + 
  geom_line(size=1.2) +
  scale_x_continuous(breaks = seq(0,20000,4000)) + # change breaks
  labs(x="N Cells in Reference", y="Running Time (Minutes)", fill="") + 
  my_theme_ggplot + RotatedAxis()

my_output_filename = paste0(my_output_prefix_figure,one_group,"_", 
                            "linechart_multiple_runs_running_time", ".png")

png(my_output_filename,
    width = 4*2*0.8, height = 3*2*0.8, units = "in",res = 300)
p
dev.off()
```

```{r, echo = FALSE, out.width='60%', fig.align='center'}
knitr::include_graphics(my_output_filename)
```

<!-- %                                                 % -->
<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

\pagebreak


