
<!-- Author: Jinming Cheng -->
<!-- Date: 17 July 2025 -->

# Prepare reference data for cell type annotation

Single-cell references for Azimuth and RCTD need to be generated and saved to disk. (References for other methods don't need to be saved to disk.)

## Reference (singlets)

Reference using my cell type annotation and singlets.

```{r,include=FALSE}
eval_prepare_azimuth_ref = FALSE
eval_prepare_RCTD_reference = FALSE
```


```{r,include=FALSE}
load("results/00_prepare.rdata")
```

Load packages
```{r loadPackage}
library(Seurat)
library(ggplot2)
library(edgeR)
library(SingleR)
library(Azimuth)
```


### Azimuth reference (Flex_Sample1)

azimuth reference dir
```{r outputPrefix}
ref_azimuth_dir = paste0(my_output_prefix,"ref_azimuth/Flex_Sample1")
if(!dir.exists(ref_azimuth_dir)){dir.create(ref_azimuth_dir,recursive = TRUE)}
```


read in scRNASeq reference data
```{r}
filename_rds = paste0(rds_res_dir,"/", "data_seurat_cell_type","_", "Flex_Sample1",".rds")
data_seurat = readRDS(file = filename_rds)
```


subset singlets as reference
```{r}
data_seurat_singlet = data_seurat[,data_seurat$db_type=="singlet"]
```

Run umap with reurn.model and normalize data using SCT method for azimuth
```{r,results='hide',eval=eval_prepare_azimuth_ref}
data_seurat_singlet = RunUMAP(data_seurat_singlet, dims=1:30,return.model = TRUE)
data_seurat_singlet = SCTransform(data_seurat_singlet)
```

Prepare azimuth reference data
```{r,results='hide',eval=eval_prepare_azimuth_ref}
ref_azimuth = AzimuthReference(data_seurat_singlet,
                               refUMAP = "umap",
                               refDR = "pca",
                               metadata = c("cell_type_label") )

saveRDS(ref_azimuth,file = paste0(ref_azimuth_dir,"/","ref.Rds"))
SaveAnnoyIndex(ref_azimuth[["refdr.annoy.neighbors"]], 
               file = paste0(ref_azimuth_dir,"/","idx.annoy"))

```



### RCTD reference (Flex_Sample1)

```{r}
library(spacexr)
```

RCTD reference dir
```{r outputPrefix}
ref_RCTD_dir = paste0(my_output_prefix,"ref_RCTD/Flex_Sample1")
if(!dir.exists(ref_RCTD_dir)){dir.create(ref_RCTD_dir,recursive = TRUE)}
```


Prepare reference
```{r,results='hide',eval=eval_prepare_RCTD_reference}
counts <- GetAssayData(data_seurat_singlet, assay = "RNA", slot = "counts")
cell_type <- as.factor(data_seurat_singlet$cell_type_label)
names(cell_type) <- colnames(data_seurat_singlet)
nUMI <- data_seurat_singlet$nCount_RNA
ref_RCTD <- Reference(counts, cell_type, nUMI)
```

save reference
```{r,eval=eval_prepare_RCTD_reference}
saveRDS(ref_RCTD,file = paste0(ref_RCTD_dir,"/","reference.rds"))
```

\pagebreak

